Response to the defense feedback

### 2021-06-02 response.txt

b. I revised Table 5.3. The evaluation section does not compare Quincunx with Abacus
in the current version because they are calculated with different methods and have
different threat models.

5. iid assumption – does it fit your research on side-channel?

It is true that Quincunx requires the assumption that input variables are
independent and identically distributed. That's why I applied a blackbox fuzzing 
instead of a coverage-guided fuzzing. As long as each time we get a random input from 
the whole input space (Population) that is independent of the previous input, the input
variables should satisfy the i.i.d. assumption. Take AES-128 encryption for example, 
each time we randomly generate a 128 bit key from $0$ to $2^{128}$, the sequence of key 
values follows the i.i.d. assumption.

The situation is a little different for a machine learning library. In my
dissertation, we also evaluated Quincunx on a deep learning library. Each time, we
randomly select an item from a dataset (Sample). If the dataset is biased, then we
can not get correct results. The good news is that the most well-known data sets in
machine learning are i.i.d. i.i.d. is the foundation of many machine learning
techniques (e.g., cross-validation, SVM, back-propagation). The assumption of i.i.d. 
simplify many methods by assuming the data won't change over the time. 
In my dissertation, we used MINIST, a well-known handwriting digit dataset. 
We believe the i.i.d. assumption holds under the circumstance, although the
assumption can be violated in some real-world scenarios.

### 2021-06-02 response.txt

Thank you for reviewing my dissertation. I have gone through the details of each
comment and revised my dissertation multiple rounds to eliminate any logic flows and
fix all the typos and grammar errors. In summary, the revision covers the following aspects.

1. Fuzzing – Why did you use fuzzing?

Abacus identifies and quantifies side-channel leakages in a single trace. So it
suffers from the problem of code coverage.  That is, Abacus misses side-channel
leakages in the unexecuted code path.

I have a few choices to address the limitation. First, I can use static methods
(verification or program analysis). Although I am interested in static analysis, I am
not a big fan of current static analysis tools, especially for side-channel research.
There are many papers on static side-channel analysis. Nearly all of these tools are
used to prove the absence of side-channel leakages. However, most of them have high
false-positive rates. Besides, there are many talented researchers working on the 
same direction (e.g., Project Everest).

On the other hand, there are many success stories with dynamic tools (e.g., SAGE,
AFL, Valgrind) in recent years. In particular, fuzzing is becoming a hot trend these
years, especially after DARPA CGC in 2016. The fuzzing method is suitable for a task
such as side-channel detections. If a program executes two different code paths or
accesses two different variables with two different input values, the program is
vulnerable to side-channel attacks. When we report side-channel leakages sites to
developers, we can tell them the two inputs. Developers can quickly reproduce the
leakage sites with the inputs.  Interestingly, there are a few (not many) related
papers. That's why I used fuzzing at the beginning of the project.

@@

It is possible to extend the quantification part to include other side-channel
signals (e.g., Sound, CPU usages, EM signals). The key is to model the noises
properly. In my dissertation, we assume that an attacker can have a noise-free
observation. While a noise-free observation is possible for some address-based
side-channel attacks, the assumption is too ideal for many other side-channel
attacks. The good news is that there are some mature methods in information theory to
model noises (e.g., Gaussian noise). The noisy-channel coding theorem provides a
computable method to estimate the information that can be reliably transferred over a
channel between secret data and attackers' observations. From a research project
aspect, I hope to have a stronger motivation before starting the project.

4.  bits vs. Q bits. Lower bounds?       Improve the presentation
    Make the comparison more meaningful
    Etc.

It is hard to explain Entropy. There is a rumor about entropy between John von
Neumann and Claude Shannon. When Shannon developed the concept of information theory,
he did not have a good idea to name the measurement of uncertainty in a system.
Neumann told Shannon, ``You should call it entropy, for two reasons. In the first
place, your uncertainty function has been used in statistical mechanics under that
name, so it already has a name. In the second place, and more important, nobody knows
what entropy really is, so in a debate, you will always have the advantage.'' During
my dissertation proposal, everyone does not really understand how we quantify the
amount of leaked information in Abacus. In my final oral exam, everyone is confused
with the term ``lower bounds'' except my advisor.

a. Abacus quantifies the amount of leaked information for a given address access
behavior. It is a measure of the amount of leaked information that an attacker can
get, based on that another event has already occurred. I agree that the result from the tool 
can be counterintuitive (e.g., the password checker example in my dissertation).
Let's compare the quantification method in my dissertation to Bayes' theorem.
I think they are similar in some ways. Bayes' theorem describes the probability of 
an event, based on prior knowledge of an event has happened. The result from Bayes' theorem 
can also be counterintuitive (e.g. drug testing).

b. Quincunx defines the amount of leaked information by the empirical entropy of the
observations. For a determinism program, channel capacity is the tight upper bound of
the empirical entropy of the observation. In the early draft of the paper, I use the
term ``conservative''. Dr. Wu thinks ``conservative'' might be a confusing term and
we decided to use the term ``lower bound''. It turns out that the term ``lower bound''
is misleading and in some cases incorrect. I have removed the term ``lower bound'' to
avoid any confusion in my dissertation.

c. I agree with Dr. Liu's comments that the quantification metrics in Chapter 4 and
Chapter 5 are different. The reason that I compared them together in Table 5.3 is that
they have the same dimensional unit (bit) under the dimensional analysis. I have
removed the comparison from the evaluation section in the current version.

To address the above concerns, I made the following revisions.

@@

a. I revised the description in Chapter 5. I avoided the use of term ``lower bound''
and tune my presentation.

b. I revised Table 5.3. The evaluation section does not compare Quincunx with Abacus
in the current version because they are calculated with different methods and have
different threat models.

c. Chapter 5 was an early draft. We submitted the work to a conference later and
my advisor helped me revise the paper. I updated the chapter from the submitted 
paper.

### 2021-06-02 response.txt

2. Discuss the limitations of your research e.g., can only handle address based …
you might want to include some discussion on the points (3), (4), and (5) below.
The research presented in the dissertation has a few limitations.

a. The methods in my dissertation can not find all kinds of side-channel leakages. No
tool can find all forms of leakages. Even for address-based side-channel leakages,
Abacus and Quincunx can not find all the leakages due to the code coverage problem.
Some verification tools can prove the absence of side-channel leakages for some small
code snippets. However, to the best of my knowledge, no tools as yet can prove the
absence of side-channel leakages in production software.

b. The quantification result only applies to specific threat models. For example,
Abacus quantifies the number of leaked bits during one real execution. So it is
possible that one leakage site leaks little information in one execution but leaks
much information in another execution.

c. The research in the dissertation can only handle address-based side-channel
attacks. Many side-channel attacks infer secret data based on other side-channel
signals (e.g., timing, EM signals). However, the root cause of these attacks is still
the same. That is, the program accesses different addresses when it processes
different input secrets. Under the circumstance, our methods can still detect these
leakage sites.

@@

It is possible to extend the quantification part to include other side-channel
signals (e.g., Sound, CPU usages, EM signals). The key is to model the noises
properly. In my dissertation, we assume that an attacker can have a noise-free
observation. While a noise-free observation is possible for some address-based
side-channel attacks, the assumption is too ideal for many other side-channel
attacks. The good news is that there are some mature methods in information theory to
model noises (e.g., Gaussian noise). The noisy-channel coding theorem provides a
computable method to estimate the information that can be reliably transferred over a
channel between secret data and attackers' observations. From a research project
aspect, I hope to have a stronger motivation before starting the project.

4.  bits vs. Q bits. Lower bounds?       Improve the presentation
    Make the comparison more meaningful
    Etc.

It is hard to explain Entropy. There is a rumor about entropy between John von
Neumann and Claude Shannon. When Shannon developed the concept of information theory,
he did not have a good idea to name the measurement of uncertainty in a system.
Neumann told Shannon, ``You should call it entropy, for two reasons. In the first
place, your uncertainty function has been used in statistical mechanics under that
name, so it already has a name. In the second place, and more important, nobody knows
what entropy really is, so in a debate, you will always have the advantage.'' During
my dissertation proposal, everyone does not really understand how we quantify the
amount of leaked information in Abacus. In my final oral exam, everyone is confused
with the term ``lower bounds'' except my advisor.

a. Abacus quantifies the amount of leaked information for a given address access
behavior. It is a measure of the amount of leaked information that an attacker can
get, based on that another event has already occurred. I agree that the result from the tool 
can be counterintuitive (e.g., the password checker example in my dissertation).
Let's compare the quantification method in my dissertation to Bayes' theorem.
I think they are similar in some ways. Bayes' theorem describes the probability of 
an event, based on prior knowledge of an event has happened. The result from Bayes' theorem 
can also be counterintuitive (e.g., drug testing).

@@

b. I revised Table 5.3. The evaluation section does not compare Quincunx with Abacus
in the current version because they are calculated with different methods and have
different threat models.

c. Chapter 5 was an early draft. We submitted the work to a conference later and
my advisor helped me revise the paper. I updated the chapter from the submitted 
paper.

5. iid assumption – does it fit your research on side-channel?

It is true that Quincunx requires the assumption that input variables are
independent and identically distributed. That's why I applied a blackbox fuzzing 
instead of a coverage-guided fuzzing. As long as each time we get a random input from 
the whole input space (Population) that is independent of the previous input, the input
variables should satisfy the i.i.d. assumption. Take AES-128 encryption for example, 
each time we randomly generate a 128 bit key from 0 to 2^{128}, the sequence of key 
values follows the i.i.d. assumption.

### 2021-06-02 response.txt

Thank you for reviewing my dissertation. I have gone through the details of each
comment and revised my dissertation multiple rounds to eliminate any logic flows and
fix all the typos and grammar errors. In summary, the revision covers the following aspects.

1. Fuzzing – Why did you use fuzzing?

Abacus identifies and quantifies side-channel leakages in a single trace. So it
suffers from the problem of code coverage. That is, Abacus misses side-channel
leakages in the unexecuted code path.

I have a few choices to address the limitation. First, I can use static methods
(verification or program analysis). Although I am interested in the static analysis, I am
not a big fan of current static analysis tools, especially for side-channel research.
There are many papers on static side-channel analysis. Nearly all of these tools are
used to prove the absence of side-channel leakages. However, most of them have high
false-positive rates. Besides, there are many talented researchers working on the 
same direction (e.g., Project Everest).

@@

b. I revised Table 5.3. The evaluation section does not compare Quincunx with Abacus
in the current version because they are calculated with different methods and have
different threat models.

c. Chapter 5 was an early draft. We submitted the work to a conference later and
my advisor helped me revise the paper. I updated the chapter from the submitted 
paper.

5. iid assumption – does it fit your research on side-channel?

It is true that Quincunx requires the assumption that input variables are
independent and identically distributed. That's why I applied a black box fuzzing 
instead of a coverage-guided fuzzing. As long as each time we get a random input from 
the whole input space (Population) that is independent of the previous input, the input
variables should satisfy the i.i.d. assumption. Take AES-128 encryption for example, 
each time we randomly generate a 128 bit key from 0 to 2^{128}, the sequence of key 
values follows the i.i.d. assumption.

The situation is a little different for a machine learning library. In my
dissertation, we also evaluated Quincunx on a deep learning library. Each time, we
randomly select an item from a dataset (Sample). If the dataset is biased, then we
can not get correct results. The good news is that the most well-known data sets in
machine learning are i.i.d. i.i.d. is the foundation of many machine learning
techniques (e.g., cross-validation, SVM, back-propagation). The assumption of i.i.d. 
simplifies many methods by assuming the data won't change over the time. 
In my dissertation, we used MINIST, a well-known handwriting digit dataset. 
We believe the i.i.d. assumption holds under the circumstance, although the
assumption can be violated in some real-world scenarios.

### 2021-06-02 Chapter-6/Chapter-6.tex
Updates from Overleaf

\subsection*{Design Choices} At the early stage of the project, we use SMT solvers to get the solutions of constraints. However, we find for some ciphers (RSA), \detect{} times out after several days. The profiling result shows that the tool spends most its time in Z3. Later, we use a sampling method to detect side-channel leakages. The sampling method seems simple and may miss some leakages in theory. However, the evaluation result shows \detect{} can identify all leakages found by the previous work~\cite{203878,236338,Brotzman19Casym}.

\subsection*{Cache Model} We use a simple cache model in this work. Recent works such as CacheAudit and CaSym~\cite{Brotzman19Casym,182946} use a Least Recently Used (LRU) cache. First, there is no fundamental difficulties for us to adopt a more realistic cache models. Second, the LRU model is still a substantial simplification of a real cache. For example, Intel Sandy Bridge CPUs use an undisclosed hash function to map memory addresses to L3 cache lines~\cite{farshin2019make}. However, the choice of LRU  cache models can bring expensive overheads. For example, CaSym can only analyze small code snippets.

\subsection*{Leakages} While recent works have reported lots of potential side-channel vulnerabilities, most of them are not patched by developers. The DES implementation of OpenSSL has a few side-channel leakages, but given the end life status of DES, it is still unpatched for the worth of engineering effort. \detect{} also detects several leakages in asymmetric ciphers such as RSA. After we manually analyze these leakage sites, we find many of them leak very little or useless information (e.g., the length of big numbers), which may partially explain why developers are not interested in fix all the side-channel leakages.

@@

\ctool{} requires the random input variables to be independent and identically distributed (i.i.d.). That's why I apply a black box fuzzing instead of a coverage-guided fuzzing. As long as each time we get a random input from the whole input space (Population) that is independent of the previous input, the input variables should satisfy the i.i.d. assumption. For AES-128 encryption, each time we randomly generate a 128 bit key from $0$ to $2^{128}$, the sequence of key values follows the i.i.d. assumption. The situation is a little different for machine learning libraries. In the dissertation, we evaluated \ctool{} on a deep learning model. Each time, we randomly select an item from a dataset (Sample). If the dataset is biased, then we can not get correct results. The good news is that the most well-known dataset in
machine learning is i.i.d. i.i.d. is the foundation of many machine learning
techniques (e.g., cross-validation, SVM, back-propagation). The assumption of i.i.d. 
simplify many methods by assuming the data won't change over the time. 
In my dissertation, we used MINIST, a well-known handwriting digit dataset. 
We believe the i.i.d. assumption holds under the circumstance, although the
assumption can be violated in some real-world scenarios.

>>>>

\subsection*{Design Choices} At the early stage of the project, we use SMT solvers to get the solutions of constraints. However, we find for some ciphers (RSA), \detect{} times out after several days. The profiling result showed that the tool spent most its time in Z3. Later, we used a sampling method to detect side-channel leakages. The sampling method seems simple and may miss some leakages in theory. However, the evaluation result shows \detect{} can identify all leakages found by the previous work~\cite{203878,236338,Brotzman19Casym}.

\subsection*{Cache Model} We use a simple cache model in this work. Recent works such as CacheAudit and CaSym~\cite{Brotzman19Casym,182946} use a Least Recently Used (LRU) cache. First, there is no fundamental difficulties that prevent us from adopting a more realistic cache models. Second, the LRU model is still a substantial simplification of a real cache. For example, Intel Sandy Bridge CPUs use an undisclosed hash function to map memory addresses to L3 cache lines~\cite{farshin2019make}. However, the choice of LRU  cache models can bring expensive overheads. For example, CaSym can only analyze small code snippets.

\subsection*{Leakages} While recent works have reported lots of potential side-channel vulnerabilities, most of them are not patched by developers. The DES implementation of OpenSSL has a few side-channel leakages, but given the end of life status of DES, it is still unpatched for the worth of engineering effort. \detect{} also detects several leakages in asymmetric ciphers such as RSA. After we manually analyze these leakage sites, we find many of them leak very little or useless information (e.g., the length of big numbers), which may partially explain why developers are not interested in fix all the side-channel leakages.

@@

\ctool{} requires the random input variables to be independent and identically distributed (i.i.d.). That's why I apply a black box fuzzing instead of a coverage-guided fuzzing. As long as each time we get a random input from the whole input space (Population) that is independent of the previous input, the input variables should satisfy the i.i.d. assumption. For AES-128 encryption, each time we randomly generate a 128 bit key from $0$ to $2^{128}$, the sequence of key values follows the i.i.d. assumption. The situation is a little different for machine learning libraries. In the dissertation, we evaluated \ctool{} on a deep learning model. Each time, we randomly select an item from a dataset (Sample). If the dataset is biased, then we can not get correct results. The good news is that the most well-known data sets in
machine learning are i.i.d. i.i.d. is the foundation of many machine learning
techniques (e.g., cross-validation, SVM, back-propagation). The assumption of i.i.d. 
simplify many methods by assuming the data won't change over the time. 
In my dissertation, we used MINIST, a well-known handwriting digit dataset. 
We believe the i.i.d. assumption holds under the circumstance, although the
assumption can be violated in some real-world scenarios.
